\large End-to-End Verification of Initial and Transition Properties of GR(1) Designs in SPARK

### 2020-07-10 doc/2020-SEFM/paper.tex
Paper edits

%\thanks{Supported by AFRL contract FA8650-16-C-2642 and AFOSR grant RQCOR20-35.}}
%
\titlerunning{End-to-End Verification of GR(1) Designs in SPARK}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Author 1\inst{1} \and
Author 2\inst{2} \and
Author 3\inst{3}}

%\author{Laura Humphrey\inst{1} \and
%James Hamil\inst{2} \and
%Joffrey Huguet\inst{3}}
%

% \authorrunning{L. Humphrey et al.}
\authorrunning{Author 1 et al.}

% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Institute 1 \\
\email{author1@email.com} \and
Institute 2 \\
\email{author2@email.com}  \and
Institute 3 \\
 \email{author3@email.com}}
%
%\institute{Air Force Research Laboratory, WPAFB, OH 45433, USA \\
%\email{laura.humphrey@us.af.mil} \and
%LinQuest Corp., Beavercreek, OH 45431, USA \\
%\email{james.hamil.ctr@us.af.mil}  \and
%AdaCore, F-75009 Paris, France \\
% \email{huguet@adacore.com}}
%
\maketitle              % typeset the header of the contribution
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 150--250 words.
Manually designing control logic for reactive systems is time-consuming and error-prone. 
An alternative is to automatically generate controllers using ``correct-by-construction'' synthesis approaches. 
Recently, there has been interest in synthesis from Generalized Reactivity(1) or GR(1) specifications, 
since the required computational complexity is relatively low, and several tools exist for synthesis from GR(1) specifications.
However, while these tools implement synthesis approaches that are theoretically ``correct-by-construction,'' 
errors in tool implementation can still lead to errors in synthesized controllers. 
We are therefore interested in ``end-to-end'' verification of synthesized controllers with respect to their original GR(1) specifications.
Toward this end, we have modified Salty -- a tool that produces executable software implementations of controllers 
from GR(1) specifications in a variety of programming languages -- to produce implementations in SPARK.
SPARK is both a programming language and associated set of verification tools, so it has the potential to enable the ``end-to-end'' verification we desire.
In this paper, we discuss our experience to date using SPARK to implement controllers and verify them against 
a subset of properties comprising GR(1) specifications, namely system initial and system transition properties. 
We also discuss lessons learned about how to best encode controllers synthesized from GR(1) specifications in SPARK for verification, 
examples in which verification found unexpected controller behaviors, and 
caveats related to the interpretation of GR(1) specifications.

@@

A GR(1) specification $\varphi$ takes the form $\varphi = \varphi^e \limplies \varphi^s$, 
where $\varphi^e$ encodes assumptions about the \emph{environment} in which a system is to operate,
and $\varphi^s$ encodes guarantees the \emph{system} should make under those assumptions \cite{Ehlers2016}. 
More specifically, $\varphi$ takes the form 
$\varphi = (\varphi^e_i \land \varphi^e_t \land \varphi^e_l)  \limplies (\varphi^s_i \land \varphi^s_t \land \varphi^s_l)$,
where $\varphi^e_i$ and $\varphi^s_i$ are \emph{initial} properties, 
$\varphi^e_t$ and $\varphi^s_t$ are \emph{transition} or safety properties, and 
$\varphi^e_l$ and $\varphi^s_l$ are \emph{liveness} properties. 
For inputs in the set $\inputs$ controlled by the environment 
and outputs in the set $\outputs$ produced by the system, terms are defined as:

@@

\noindent 
In this context, the temporal operators $\always$ ``always,'' $\eventually$ ``eventually,'' and $\lnext$  ``next'' have the following meanings. 
A formula of the form $\always b$ holds if $b$ is true at every time step, 
$\eventually b$ holds if $b$ is eventually true at some future time step, 
$\always \eventually b$ holds if $b$ is true infinitely often in the future, 
and $\lnext b$ holds if $b$ is true at the next time step. 
It is assumed that at each time step, the environment chooses an input from $\inputs$, then the system chooses an output from $\outputs$ in response. 
The result of synthesis is a \emph{control protocol} or \emph{strategy} that can be expressed as a Moore machine. 
%Note that in the case of GR(1), synthesized Moore machines not only produce a single set of system output values in each state, 
%there is a unique set of environment input values that brings the controller into each state.

@@

SPARK, which is based on the Ada programming language, is both a programming language 
with a specification language and an associated verification toolset \cite{hoang2015spark}.
% SPARK translates code to the Why language for processing by the Why3 theorem prover, 
% and Why3 uses other automated provers -- including SMT solvers Z3, CVC4, and Alt-Ergo -- to discharge verification conditions.
Though SPARK aims to perform fully automated verification, manual adaptation of the source code is in general necessary. 
For example, if one were to translate source code originally written in Ada to SPARK, 
one would have to remove unsupported features such as functions with side-effects, aliased variable names, exception handlers, etc.  
Formal verification of user-specified properties also requires manually writing specifications at the level of the source code in the 
form of contracts, e.g. preconditions and postconditions that summarize the assumptions and guarantees provided by individual subprograms.
There are different levels to which one can use SPARK to verify a program \cite{moy2019climbing}. 
Generally, the level of verification performed during development is incremental, going from lowest to highest. 
Verification at the lowest level, colloquially referred to as ``stone level,'' is achieved when code is accepted by SPARK, 
since the SPARK language has stricter legality rules than Ada. 
Verification at the next level, ``bronze level,'' is achieved when flow analysis returns with no error; 
flow analysis checks for errors such as reads of uninitialized data or violations of user-specified data flow contracts. 
Verification at the next level, ``silver level,'' ensures that no runtime errors will be encountered when executing the program, 
e.g. division by zero or numeric overflow/underflow. 
Verification at the next level, ``gold level,''  consists of verifying key user specifications, such as 
type invariants and subprogram preconditions and postconditions; 
however, at this level, the specifications only partially specify the desired behavior of the code.  
Verification at the next level, ``platinum level,'' consists of verifying a complete set of specifications.
At the moment, the SPARK code generated by Salty proves at gold level.
In particular, SPARK is able to prove the system initial and system transition portions the original GR(1) specification. 
Proof of system transition properties in turn requires proving a type invariant on the underlying Moore machine representation of the controller; 
this type invariant encodes that each state is reached by a specific set of input values and produces a specific set of output values. 
Platinum level would be reached if the liveness properties were expressed and proved, 
which we leave for future work.

In what follows, in \sectref{sec:implementation} we describe the structure of synthesized controller implementations in SPARK 
by walking through a simple example involving a traffic light. 
To evaluate whether verification of synthesized SPARK controllers is feasible and has utility, 
we collected a corpus of examples from various sources. 
In \sectref{sec:vipExample}, we take a detailed look at one of these examples, a controller that coordinates the actions of 
a team of unmanned air vehicles (UAVs) performing an escort mission, 
and we describe how SPARK revealed an error in the controller's specification. 
In \sectref{sec:caseStudies}, we give an overview of results for the rest of the examples, with a focus on scalability. 
In \sectref{sec:lessonsLearned}, we discuss lessons learned, including how to best structure synthesized controllers for proof.  
% examples in which SPARK revealed unexpected controller behaviors, and caveats related to GR(1) specifications. 
We end with concluding remarks in \sectref{sec:conclusions}, including a discussion of future work.

@@

In SPARK, subunits consist of a specification and a body.
% For each synthesized controller, Salty automatically generates both a specification and body for a package named according to the Salty specification. 
The package specification for this traffic light example is shown in \figref{fig:trafficLightSpec}.
Type \lstinline{Controller} encodes a Moore machine representing the synthesized controller. 
In the public part of the specification, \lstinline{Controller} (line 2) is declared as a private type so that the user cannot arbitrarily manipulate its state.  
% (possibly violating the controller's GR(1) specification). 
In the private part of the specification, \lstinline{Controller} (lines 33-37) is a record that stores the controller's internal state number, 
current input value(s), and current output value(s). 
The internal state number is always initialized to the last possible state number; 
this ``controller initialization'' state encodes the status of the controller before any inputs are received, and the only transition(s) out of this state are those allowed by $\varphi_i^e$.
In this example, each input and output is of type \lstinline{Boolean} (Salty also allows for enumerations and integers, to be discussed later). 
When there are multiple inputs or outputs, they are wrapped in a record of type \lstinline{Environment} or \lstinline{System} (lines 4-6), respectively. 
Here, there is only one input, so it is not wrapped in a record. 
Since \lstinline{Controller} is a Moore machine, for each internal state, 
there is exactly one set of output values produced in that state. 
Furthermore, recall that controllers synthesized from GR(1) specifications have the additional property that 
there is exactly one set of input values that brings the controller into each state. 
\lstinline{Controller} therefore includes a type invariant (lines 36-37) that captures this property, 
where functions \lstinline{State_To_Input_Mapping} and \lstinline{State_To_Output_Mapping} evaluate 
whether the controller's state corresponds to the expected input and output value(s), respectively. 
% They use lookup tables, \lstinline{State_To_Input_Equivalence} and \lstinline{State_To_Output_Equivalence}, defined in the package body, to do so.
Both of these functions are declared (lines 26-29) with aspect \lstinline{Ghost}, indicating that they are intended mainly for proof purposes, 
i.e. they will create verification conditions related to type \lstinline{Controller} but will not be executed in the actual program, 
unless it is specified during compilation that they should be.
Executing ghost code is often used for debugging, e.g. to check an unproved property through testing. 
In this case, removing the ghost code can save significant memory, 
since \lstinline{State_To_Input_Mapping} and \lstinline{State_To_Output_Mapping} 
internally encode large lookup tables that have to store information on every state of the controller, 
which is often thousands of states or even millions of states for the largest example in our database.
The logic for these functions is given in the body (not shown here), but for instance when \lstinline{C.State = 1}, 
\lstinline{State_To_Input_Mapping(C)} returns \lstinline{True} if and only if \lstinline{C.tick = False}, and 
\lstinline{State_To_Output_Mapping(C)} returns \lstinline{True} if and only if 
\sloppy{\lstinline{C.S = System'(red => True, yellow => False, green => False)}}, 
since \lstinline{tick = False} brings the system into state 1, and $red$ is true and $yellow$ and $green$ are false in state 1.

@@

In order to evaluate the feasibility and utility of our approach, we collected GR(1) specifications from a variety of sources. 
This includes the Salty repository, which contains examples of GR(1) specifications for control of teams of UAVs performing different missions. 
One of these encodes the rules for a ``VIP Escort'' mission, 
in which one UAV is designated as a ``very important person'' (VIP) that must always be ``escorted'' by a friendly surveillance UAV when it moves, 
and it must also be protected from an ``enemy'' UAV. 
The VIP and surveillance UAVs are controlled by the system, while the enemy UAV is controlled by the environment.
The mission map contains regions that the UAVs can move between.  
``Escorting'' consists of ensuring that 1) the VIP only enters regions previously visited by a surveillance UAV 
and 2) whenever the VIP changes regions, a surveillance UAV ``tracks'' it, i.e. moves with it between regions. 
``Protection'' consists of ensuring that the VIP is never in the same region as the enemy UAV. 
Additional rules include constraints and liveness requirements on how the UAVs can move between regions and which regions they start in.

@@

This error is the result of a subtlety of the semantics of GR(1) specifications. 
Recall that GR(1) specifications take the form $\varphi_e \limplies \varphi_s$. 
Obviously a specification of this form is satisfied if $\varphi_e$ and $\varphi_s$ are both true, 
but it is also satisfied if $\varphi_e$ is false. 
Also recall from the introduction that GR(1) specifications are interpreted in the context of a two-player game 
in which the environment takes its turn first and the system takes its turn second. 
The issue here is that the environment is able to take transitions that will necessarily cause it to violate $\varphi_e^t$ in the next time step. 
Note that $\varphi_e^t$ contains terms of the form $\always \lnot p$, specifically $\always \lnot (loc_e = i)$ for $i = \{1, 2\}$  \eqnref{eq:vipEnvDisallowedRegions}. 
Note that a term of the form $\always \lnot p$ is not the same as a term of the form $\always \lnext \lnot p$. 
The distinction is important. 
If the environment chooses $p$ for the ``next'' time step, this does not violate $\always \lnot p$ in the ``current'' time step. %, since the new ``current'' time step is only constructed once the environment and system have both taken their turn. 
However, once the next state is reached, $p$ becomes the new ``current'' value, and $\always \lnot p$ 
will now be violated no matter what the environment chooses.
Generally, specifications should follow the latter form $\always \lnext \lnot p$, 
which prohibits the environment from choosing $p$ in the ``next'' time step. 
This was indeed an error in these specifications, so we changed $\always \lnot (loc_e = 1) \land \always \lnot (loc_e = 2)$
 to $\always \lnext \lnot (loc_e = 1) \land \always \lnext \lnot (loc_e = 2)$ in \eqnref{eq:vipEnvDisallowedRegions}.
However, we also modified Salty to \lstinline{raise Program_Error} in cases with no successors 
and checked that we were still able to prove $\varphi_e \limplies \varphi_s$ 
(since reaching these cases would require violating the precondition $\varphi_e$). 
Such cases amount to instances in which the precondition on \lstinline{Move} is would have to be violated, which is why we allow the user to execute \lstinline{Env_Trans} 
as discussed in \sectref{sec:implementation}. 
We also plan to have Salty issue a warning when there are states with no successors, 
since such cases are likely unintended by the user.

>>>>

%\thanks{Supported by AFRL contract FA8650-16-C-2642 and AFOSR grant RQCOR20-35.}}
%
\titlerunning{End-to-End Verification of GR(1) Designs in SPARK}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
% \author{Author 1\inst{1} \and
% Author 2\inst{2} \and
% Author 3\inst{3}}

\author{Laura R. Humphrey\inst{1} \and
James Hamil\inst{2} \and
Joffrey Huguet\inst{3}}
%

\authorrunning{L. Humphrey et al.}

% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Air Force Research Laboratory, WPAFB, OH 45433, USA \\
\email{laura.humphrey@us.af.mil} \and
LinQuest Corp., Beavercreek, OH 45431, USA \\
\email{james.hamil.ctr@us.af.mil}  \and
AdaCore, F-75009 Paris, France \\
 \email{huguet@adacore.com}}
%
\maketitle              % typeset the header of the contribution
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 150--250 words.
Manually designing control logic for reactive systems is time-consuming and error-prone. 
An alternative is to automatically generate controllers using ``correct-by-construction'' synthesis approaches. 
Recently, there has been interest in synthesis from Generalized Reactivity(1) or GR(1) specifications, 
since the required computational complexity is relatively low, and several tools exist for synthesis from GR(1) specifications.
However, while these tools implement synthesis approaches that are theoretically ``correct-by-construction,'' 
errors in tool implementation can still lead to errors in synthesized controllers. 
We are therefore interested in ``end-to-end'' verification of synthesized controllers with respect to their original GR(1) specifications.
Toward this end, we have modified Salty -- a tool that produces executable software implementations of controllers 
from GR(1) specifications in a variety of programming languages -- to produce implementations in SPARK.
SPARK is both a programming language and associated set of verification tools, so it has the potential to enable the ``end-to-end'' verification we desire.
In this paper, we discuss our experience to date using SPARK to implement controllers and verify them against 
a subset of properties comprising GR(1) specifications, namely system initial and system transition properties. 
We also discuss lessons learned about how to best encode controllers synthesized from GR(1) specifications in SPARK for verification, 
examples in which verification found unexpected controller behaviors, and 
caveats related to the interpretation of GR(1) specifications.

@@

GR(1) is a fragment of linear temporal logic. 
A GR(1) specification $\varphi$ takes the form $\varphi = \varphi^e \limplies \varphi^s$, 
where $\varphi^e$ encodes assumptions about the \emph{environment} in which a system is to operate,
and $\varphi^s$ encodes guarantees the \emph{system} should make under those assumptions \cite{Ehlers2016}. 
More specifically, $\varphi$ takes the form 
$\varphi = (\varphi^e_i \land \varphi^e_t \land \varphi^e_l)  \limplies (\varphi^s_i \land \varphi^s_t \land \varphi^s_l)$,
where $\varphi^e_i$ and $\varphi^s_i$ are \emph{initial} properties, 
$\varphi^e_t$ and $\varphi^s_t$ are \emph{transition} or safety properties, and 
$\varphi^e_l$ and $\varphi^s_l$ are \emph{liveness} properties. 
For inputs in the set $\inputs$ controlled by the environment 
and outputs in the set $\outputs$ produced by the system, terms are defined as:

@@

\noindent 
In this context, the temporal operators $\always$ ``always,'' $\eventually$ ``eventually,'' and $\lnext$  ``next'' have the following meanings. 
A formula of the form $\always b$ holds if $b$ is true at every time step, 
$\eventually b$ holds if $b$ is eventually true at some future time step, 
$\always \eventually b$ holds if $b$ is true infinitely often in the future, 
and $\lnext b$ holds if $b$ is true at the next time step. 
It is assumed that at each time step, the environment chooses an input from $\inputs$, then the system chooses an output from $\outputs$ in response. 
Synthesis from GR(1) specifications can therefore be viewed as a two-player game between the system and the environment, 
where the goal is for the system to satisfy $\varphi^s$ as long as the environment satisfies $\varphi^e$. 
If this goal is achievable, then we say the specification is realizable, 
and the result is a \emph{control protocol} or \emph{strategy} for the system that can be expressed as a Moore machine. 
%Note that in the case of GR(1), synthesized Moore machines not only produce a single set of system output values in each state, 
%there is a unique set of environment input values that brings the controller into each state.

@@

SPARK, which is based on the Ada programming language, is both a programming language 
with a specification language and an associated verification toolset \cite{hoang2015spark}.
% SPARK translates code to the Why language for processing by the Why3 theorem prover, 
% and Why3 uses other automated provers -- including SMT solvers Z3, CVC4, and Alt-Ergo -- to discharge verification conditions.
Though SPARK aims to perform fully automated verification, manual adaptation of the source code is in general necessary. 
For example, if one were to translate source code originally written in Ada to SPARK, 
one would have to remove unsupported features such as functions with side-effects, aliased variable names, exception handlers, etc.  
Formal verification of user-specified properties also requires manually writing specifications at the level of the source code in the 
form of contracts, e.g. preconditions and postconditions that summarize the assumptions and guarantees provided by individual subprograms.
There are different levels to which one can use SPARK to verify a program \cite{moy2019climbing}. 
Generally, the level of verification performed during development is incremental, going from lowest to highest. 
Verification at the lowest level, colloquially referred to as \emph{stone level}, is achieved when code is accepted by SPARK, 
since the SPARK language has stricter legality rules than Ada. 
Verification at the next level, \emph{bronze level}, is achieved when flow analysis returns with no error; 
flow analysis checks for errors such as reads of uninitialized data or violations of user-specified data flow contracts. 
Verification at the next level, \emph{silver level}, ensures that no runtime errors will be encountered when executing the program, 
e.g. division by zero or numeric overflow/underflow. 
Verification at the next level, \emph{gold level}, consists of verifying key user specifications, such as 
type invariants and subprogram preconditions and postconditions; 
however, at this level, the specifications only partially specify the desired behavior of the code.  
Verification at the next level, \emph{platinum level}, consists of verifying a complete set of specifications.
At the moment, the SPARK code generated by Salty proves at gold level.
In particular, SPARK is able to prove the system initial and system transition portions the original GR(1) specification. 
Proof of system transition properties in turn requires proving a type invariant on the underlying Moore machine representation of the controller; 
this type invariant encodes that each state is reached by a specific set of input values and produces a specific set of output values. 
Platinum level would be reached if the liveness properties were expressed and proved, 
which we leave for future work.

In what follows, in \sectref{sec:implementation} we describe the structure of synthesized controller implementations in SPARK 
by walking through a simple example involving a traffic light. 
To evaluate whether verification of synthesized SPARK controllers is feasible and has utility, 
we collected a corpus of examples from various sources. 
In \sectref{sec:vipExample}, we take a detailed look at one of these examples, a controller that coordinates the actions of 
a team of unmanned air vehicles performing an escort mission, 
and we describe how SPARK revealed an error in the controller's specification. 
In \sectref{sec:caseStudies}, we give an overview of results for the rest of the examples, with a focus on scalability. 
In \sectref{sec:lessonsLearned}, we discuss lessons learned, including how to best structure synthesized controllers for proof.  
% examples in which SPARK revealed unexpected controller behaviors, and caveats related to GR(1) specifications. 
We end with concluding remarks in \sectref{sec:conclusions}, including a discussion of future work.

@@

In SPARK, subunits consist of a specification and a body.
% For each synthesized controller, Salty automatically generates both a specification and body for a package named according to the Salty specification. 
The package specification for this traffic light example is shown in \figref{fig:trafficLightSpec}.
Type \lstinline{Controller} encodes a Moore machine representing the synthesized controller. 
In the public part of the specification, \lstinline{Controller} (line 2) is declared as a private type so that the user cannot arbitrarily manipulate its state.  
% (possibly violating the controller's GR(1) specification). 
In the private part of the specification, \lstinline{Controller} (lines 33-37) is a record that stores the controller's internal state number, 
current input value(s), and current output value(s). 
The internal state number is always initialized to the last possible state number, i.e. the largest value for \lstinline{State_Num}; 
this ``controller initialization'' state encodes the status of the controller before any inputs are received, and the only transition(s) out of this state are those allowed by $\varphi_i^e$.
In this example, each input and output is of type \lstinline{Boolean} (Salty also allows for enumerations and integers, to be discussed later). 
When there are multiple inputs or outputs, they are wrapped in a record of type \lstinline{Environment} (not used here) or \lstinline{System} (lines 4-6), respectively. 
There is only one input in this example, so it is not wrapped in a record, which is why there is no \lstinline{Environment} record in \figref{fig:trafficLightSpec}. 
Since \lstinline{Controller} is a Moore machine, for each internal state, 
there is exactly one set of output values produced in that state. 
Furthermore, recall that controllers synthesized from GR(1) specifications have the additional property that 
there is exactly one set of input values that brings the controller into each state. 
\lstinline{Controller} therefore includes a type invariant (lines 36-37) that captures this property, 
where functions \lstinline{State_To_Input_Mapping} and \lstinline{State_To_Output_Mapping} evaluate 
whether the controller's state corresponds to the expected input and output value(s), respectively. 
% They use lookup tables, \lstinline{State_To_Input_Equivalence} and \lstinline{State_To_Output_Equivalence}, defined in the package body, to do so.
Both of these functions are declared (lines 26-29) with aspect \lstinline{Ghost}, indicating that they are intended mainly for proof purposes, 
i.e. they will create verification conditions related to type \lstinline{Controller} but will not be executed in the actual program, 
unless it is specified during compilation that they should be.
Executing ghost code is often used for debugging, e.g. to check an unproved property through testing. 
In this case, removing the ghost code can save significant memory, 
since \lstinline{State_To_Input_Mapping} and \lstinline{State_To_Output_Mapping} 
internally encode large lookup tables that have to store information on every state of the controller, 
which is often thousands of states or even millions of states for the largest example in our database.
The logic for these functions is given in the body (not shown here), but for instance when \lstinline{C.State = 1}, 
\lstinline{State_To_Input_Mapping(C)} returns \lstinline{True} if and only if \lstinline{C.tick = False}, and 
\lstinline{State_To_Output_Mapping(C)} returns \lstinline{True} if and only if 
\sloppy{\lstinline{C.S = System'(red => True, yellow => False, green => False)}}, 
since \lstinline{tick = False} brings the system into state 1, and $red$ is true and $yellow$ and $green$ are false in state 1.

@@

In order to evaluate the feasibility and utility of our approach, we collected GR(1) specifications from a variety of sources. 
This includes the Salty repository, which contains examples of GR(1) specifications 
for control of teams of unmanned air vehicles (UAVs) performing different missions. 
One of these encodes the rules for a ``VIP Escort'' mission, 
in which one UAV is designated as a ``very important person'' (VIP) 
that must always be ``escorted'' by a friendly surveillance UAV when it moves, 
and it must also be protected from an ``enemy'' UAV. 
The VIP and surveillance UAVs are controlled by the system, while the enemy UAV is controlled by the environment.
The mission map contains regions that the UAVs can move between.  
``Escorting'' consists of ensuring that 1) the VIP only enters regions previously visited by a surveillance UAV 
and 2) whenever the VIP changes regions, a surveillance UAV ``tracks'' it, i.e. moves with it between regions. 
``Protection'' consists of ensuring that the VIP is never in the same region as the enemy UAV. 
Additional rules include constraints and liveness requirements on how the UAVs can move between regions and which regions they start in.

@@

This error is the result of a subtlety of the semantics of GR(1) specifications. 
Recall that GR(1) specifications take the form $\varphi_e \limplies \varphi_s$. 
Obviously a specification of this form is satisfied if $\varphi_e$ and $\varphi_s$ are both true, 
but it is also satisfied if $\varphi_e$ is false. 
Also recall from the introduction that GR(1) specifications are interpreted in the context of a two-player game 
in which the environment takes its turn first and the system takes its turn second. 
The issue here is that the environment is able to take transitions that will necessarily cause it to violate $\varphi_e^t$ in the next time step. 
Note that $\varphi_e^t$ contains terms of the form $\always \lnot p$, specifically $\always \lnot (loc_e = i)$ for $i = \{1, 2\}$  \eqnref{eq:vipEnvDisallowedRegions}. 
Note that a term of the form $\always \lnot p$ is not the same as a term of the form $\always \lnext \lnot p$. 
The distinction is important. 
If the environment chooses $p$ for the ``next'' time step, this does not violate $\always \lnot p$ in the ``current'' time step. %, since the new ``current'' time step is only constructed once the environment and system have both taken their turn. 
However, once the next state is reached, $p$ becomes the new ``current'' value, and $\always \lnot p$ 
will now be violated no matter what the environment chooses.
Generally, specifications should follow the latter form $\always \lnext \lnot p$, 
which prohibits the environment from choosing $p$ in the ``next'' time step. 
This was indeed an error in these specifications, so we changed $\always \lnot (loc_e = 1) \land \always \lnot (loc_e = 2)$
 to $\always \lnext \lnot (loc_e = 1) \land \always \lnext \lnot (loc_e = 2)$ in \eqnref{eq:vipEnvDisallowedRegions}.
However, we also modified Salty to \lstinline{raise Program_Error} in cases with no successors 
and checked that we were still able to prove $\varphi_e \limplies \varphi_s$ 
(since reaching these cases would require violating the precondition $\varphi_e$). 
Such cases amount to instances in which the precondition on \lstinline{Move} would have to be violated, 
which is why we allow the user to execute \lstinline{Env_Trans} 
as discussed in \sectref{sec:implementation}. 
We also plan to have Salty issue a warning when there are states with no successors, 
since such cases are likely unintended by the user.
